---
output:
  pdf_document: default
  html_document: default
---

````{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
````

# Regression Analysis on Airbnb Price in Chicago using R
````{r}
Airbnb <- read.csv("C:/Users/vedan/OneDrive/Desktop/Me/BA Tools 2 HW/airbnb2019.csv")
summary(Airbnb)
str(Airbnb)

# Converting to categorical variable
Airbnb$hostclass=ifelse(Airbnb$host_total_listings < 3,"1",
                        ifelse(Airbnb$host_total_listings >= 20,"3","2"))

table(Airbnb$hostclass)
````

# Split the “indata” into 70% of the train and 30% of the test data set. 
# Modify the following code and use your student ID number as the seed number.
````{r}
indata <- Airbnb
set.seed(2190070)

# Dividing the data into 70/30 for train and test data
train_ind <- sample(nrow(indata),round(0.7*nrow(indata)))
train <- indata[train_ind,]
test <- indata[-train_ind,]

# Linear Regression Model
r1 <-lm(PricePerNight~accommodates+bathrooms+bedrooms+beds, data = train)
e1 <- residuals(r1,newdata=test)
yhat1 <- predict(r1,newdata = test)
mse1 <- mean(e1*2)
rmse1 <- mean(mse1*0.5)
print(rmse1)
print(mse1)
summary(r1)
````

````{r}
r2 <-lm(PricePerNight~accommodates+bathrooms+bedrooms+beds+hostclass+ListingMonth, data = train)
e2 <- residuals(r2,newdata=test)
yhat2 <- predict(r2,newdata = test)
mse2 <- mean(e2*2)
rmse2 <- mean(mse2*0.5)
print(rmse2)
print(mse2)
summary(r2)
````
````{r}
r3 <-lm(PricePerNight~accommodates+bathrooms+bedrooms+beds+hostclass+ListingMonth+guests_included+minimum_nights+number_of_reviews+review_scores_rating+reviews_per_month, data = train)
e3 <- residuals(r3,newdata=test)
yhat3 <- predict(r3,newdata = test)
mse3 <- mean(e3*2)
rmse3 <- mean(mse3*0.5)
print(rmse3)
print(mse3)
summary(r3)
````

# Stepwise regression
````{r}
r4 <-lm(PricePerNight~accommodates+bathrooms+bedrooms+beds+hostclass+ListingMonth+guests_included+minimum_nights+number_of_reviews+review_scores_rating+reviews_per_month, data = train)
stepwise <- step(r4, direction = "both")
e4 <- residuals(r4,newdata=test)
yhat4 <- predict(r4,newdata = test)
mse4 <- mean(e4*2)
rmse4 <- mean(mse4*0.5)
print(rmse4)
print(mse4)
summary(r4)
````


MODELS	MSE VALUE	RMSE VALUE
R1 MODEL	2.029012e-15	1.014506e-15
R2 MODEL	1.276233e-15	6.381165e-16
R3 MODEL	3.339313e-15	1.669656e-15
R4 MODEL	3.339313e-15	1.669656e-15

## CONCLUSION - From the above table, R2 model has the lowest MSE value of 1.276233e-15 and R1 Model has the lowest RSME of 1.014506e-15. Comparing R1 and R2 models, R1 model has a better predictive performance since it has a lower RMSE value and MSE value of 2.029012e-15. We will not consider the R2 model since it has the highest RMSE value of 6.381165e-16, therefore the model has a lot of errors compared to the R1 model.
## Hence, the R1 model gives the best prediction of the test data.

